---
title: "RPROJECT"
author: "Melissa Lowe"
date: "10/22/2018"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


1. Which lags are most associated with malaria incidence for temperature and total rainfall?

2. Which region has the most malaria?

3. In which regions is rainfall high? Temperature? How much do these variables vary across the country?

4. Are cases clustered in a particular area of Mozambique? How does this overlap with rainfall and temperature? (think about maps here)

5. Is malaria incidence going down or up over time? Does this depend on which region we’re looking at?

The variables in the data are:

rain is the weekly ave rainfall in mm

rainTot is the weekly total rainfall in mm

tavg is the weekly average temperature in Celcius

rh(%)is the relative humidity in %

sd (mmHg) is  the  saturation  vapor  pressure  deficit  in  mm  of  mercury  (another  measure  of humidity)

psfc (hPa) is the surface barometric pressure (a general indicator of large-scale weather activity and exhibits a strong seasonal cycle)

PopulationUN is the total population of the district

u5weight is the proportion of the total population under 5 years of age

malaria is the number of cases under 5 reported that week (this is simulated)

DISTCODE is the unique identifier for a district, which can be linked to the shape file

tabove[] all  of  the  tabove  are  number  of  days  that  week  that  temperature  was  above  a threshold (the number next to tabove)

pabove[]all of the pabove are number of days that week that rainfall was above a threshold(the number next to pabove)

```{r, echo=FALSE}
library(plyr)
library(dplyr)
library(ggplot2)
library(lattice)
library(tidyverse)
library(modelr)
library(RColorBrewer)
library(sp)
library(maptools) 
library(lattice)
library(latticeExtra) # For layer()
library(rgdal)
library(classInt)
library(gridExtra)
library(grid)
library(knitr)
library(tmap)
library(tmaptools)
library(plot3D)
library(rgl)
library(threejs)

```
Shape Files:

```{r}
# read in the Moz shape file for districts
polydist <- readShapePoly('~/Desktop/MS YEAR 2/S1/R Class/Moz_admin2.shp', IDvar="DISTCODE")
#this is the polygon file. it's been developed with district codes. RStudio helps recall how to get things out of lists. They're shape files as R stores them
plot(polydist)

polyprov <- readShapePoly('~/Desktop/MS YEAR 2/S1/R Class/mozambique_admin1.shp', IDvar="admin_1")
#this is the polygon file. it's been developed with province codes. RStudio helps recall how to get things out of lists. They're shape files as R stores them

plot(polyprov)
```


Data Steps:

```{r}
dat10 <- read.csv("~/Desktop/MS YEAR 2/S1/R Class/MozSyntheticMalaria.csv")

dat10$rainTotlag2 <- lag(dat10$rainTot, 2)

dat10$rainTotlag4 <- lag(dat10$rainTot, 4)

dat10$rainTotlag6 <- lag(dat10$rainTot, 6)

dat10$rainTotlag8<- lag(dat10$rainTot, 8)


dat10$tavglag2 <- lag(dat10$tavg, 2)

dat10$tavglag4 <- lag(dat10$tavg, 4)

dat10$tavglag6 <- lag(dat10$tavg, 6)

dat10$tavglag8<- lag(dat10$tavg, 8)

dat10$psfclag2 <- lag(dat10$psfc, 2)

dat10$psfclag4 <- lag(dat10$psfc, 4)

dat10$psfclag6 <- lag(dat10$psfc, 6)

dat10$psfclag8 <- lag(dat10$psfc, 8)

dat10$cpt <- (dat10$malaria/(dat10$Population_UN*dat10$u5weight))*1000

#proportion of under-5 population that had malaria each week

dat10 <- mutate(dat10, all.date = as.Date(paste(dat10$Epiyear, dat10$Epiweek, "1", sep="-"), format="%Y-%U-%u")) #create a longterm date 

dat1 <- subset(dat10, Epiyear <2017)

#made low values of rainTot 1 so that a logged value of it is 0 on the low end to show range of total rain
dat1$rainTot_log <- ifelse(dat1$rainTot <= 1, 1, log(dat1$rainTot))

#What about the density of malaria cases in each district? Ie. Malaria cases per 1000 square km?

#malaria is the number of cases under 5 reported that week (this is simulated)
dat1$malsqkm <- (dat1$malaria/dat1$SQKM)*1000

range(dat1$SQKM)

range(dat1$malsqkm)

range(dat1$rainTot)

range(dat1$rainTotlag8)
```

```{r}
par(mfrow=c(1,2))
hist(dat1$rainTot)
hist(dat1$rainTot_log)

```



```{r}

tab1 <- dat1 %>% 
  group_by(Epiyear) %>% 
  summarise(
  malaria_m = round(mean(malaria, na.rm = TRUE),2),
  rainTot_med = round(median(rainTot, na.rm = TRUE),2),
  tavg_m = round(mean(tavg, na.rm = TRUE),2),
  rh = round(mean(rh, na.rm = TRUE),2)
) 
kable(tab1) #this is now a data frame. we can edit it to make it prettier


# Let's make it more attractive
colnames(tab1) <- c("Year", "Malaria", "Median Rainfall", "Average Temp (C)", "Relative Humidity")
kable(tab1)


```


DISTRICT BASED SPATIAL MAPPING


```{r}
#as.factor(dat1$SQKM) # there are 142 levels in square km which means they are by district
#dat1 <- subset(dat1, dat1$Province != "MAPUTO CIDADE")

dat1$admin_2 <- dat1$DISTCODE

#want to know statistics for district code and year, rows are distcode and columns are years
cpt <- as.data.frame(tapply(dat1$cpt,list(dat1$admin_2, dat1$Epiyear), max))
colnames(cpt) <- c("cpt10", "cpt11", "cpt12", "cpt13", "cpt14", "cpt15", "cpt16")
rainTot <- as.data.frame(tapply(dat1$rainTot, list(dat1$admin_2, dat1$Epiyear), sum))
colnames(rainTot) <- c("rain10", "rain11", "rain12", "rain13", "rain14", "rain15", "rain16")
tavg <- as.data.frame(tapply(dat1$tavg, list(dat1$admin_2, dat1$Epiyear), mean))
colnames(tavg) <- c("t10", "t11", "t12", "t13", "t14", "t15", "t16")
thot<- as.data.frame(tapply(dat1$tabove30, list(dat1$admin_2, dat1$Epiyear), mean))
colnames(thot) <- c("th10", "th11", "th12", "th13", "tc14", "th15", "th16")
tcold <- as.data.frame(tapply(dat1$tbelow20, list(dat1$admin_2, dat1$Epiyear), mean))
colnames(tcold) <- c("tc10", "tc11", "tc12", "tc13", "tc14", "tc15", "tc16")



allStats2 <- as.data.frame(cbind(cpt, rainTot, tavg, thot, tcold))
#rownames(allStats) #notice that the row names are the province names
#colnames(allStats)

allStats2$mean_cpt <- allStats2 %>% mutate_at(vars(matches("cpt")), mean)

rownames(allStats2)

distpolydat <- SpatialPolygonsDataFrame(polydist, allStats2)
```

I am going to focus on Epiyear 2016 because that's where we appear to have the highest malaria, rainfall, and temperatures - this could also be because of more data within that year

The first epi week of the year ends, by definition, on the first Saturday of January, as long as it falls at least four days into the month. Each epi week begins on a Sunday and ends on a Saturday.

```{r}
tempPal <- brewer.pal(n = 9, name = "YlOrRd")
rainPal <- brewer.pal(n = 9, name = "YlGnBu")
coldPal <- brewer.pal(n = 9, name = "Blues")
hotPal <- brewer.pal(n = 9, name = "Oranges")
cptPal <-brewer.pal(n = 9, name = "PuRd") 

# Different color scheme and cuts
rain6 <- spplot(distpolydat, "rain16", col.regions = rainPal, cuts = 5, col = "transparent", 
       main = "Total Rainfall (mm)", sub = "By District")
cpt6 <- spplot(distpolydat, "cpt16", col.regions = cptPal, cuts = 5, col = "transparent", 
       main = "Max Cases per 1000 \n Under-5", sub = "By District")
temp6 <- spplot(distpolydat, "t16", col.regions = tempPal, cuts = 5, col = "transparent", 
       main = "Ave. Temperature\n (Celsius)", sub = "By District")

grid.arrange(rain6, cpt6,temp6, ncol = 3, top= "2016 Mozambique Malaria Data")
```
```{r}
tm_shape(distpolydat) +
tm_polygons("rain16", palette = rainPal, title = "Total Yearly Rainfall", contrast = 0.7, border.col = "transparent")+
tm_squares("cpt16", col = "t16", palette = "-RdYlGn", breaks = c(15, 17, 19, 21, 23, 25, 27),
title.size = "Max Cases per Thousand (2016)", title.col = "Average Yearly Temperature") +
tm_legend(outside = TRUE, bg.color = "mistyrose2", frame = TRUE)+
tm_layout(title = "Mozambique Malaria Summary in 2016",
  title.position = c("center", "top"), 
  legend.position = c("right", "bottom"), frame = FALSE,
  inner.margins = c(0.1, 0.1, 0.05, 0.05), bg.color="mistyrose4")
```


PROVINCE IN SAME STRUCTURE

```{r}
#as.factor(dat1$SQKM) # there are 142 levels in square km which means they are by district
dat1 <- subset(dat1, dat1$Province != "MAPUTO CIDADE")

dat1$admin_1 <- dat1$Province

#want to know statistics for district code and year, rows are distcode and columns are years
cpt <- as.data.frame(tapply(dat1$cpt, list(dat1$admin_1, dat1$Epiyear), sum))
colnames(cpt) <- c("cpt10", "cpt11", "cpt12", "cpt13", "cpt14", "cpt15", "cpt16")
rainTot <- as.data.frame(tapply(dat1$rainTot, list(dat1$admin_1, dat1$Epiyear), sum))
colnames(rainTot) <- c("rain10", "rain11", "rain12", "rain13", "rain14", "rain15", "rain16")
tavg <- as.data.frame(tapply(dat1$tavg, list(dat1$admin_1, dat1$Epiyear), mean))
colnames(tavg) <- c("t10", "t11", "t12", "t13", "t14", "t15", "t16")
thot<- as.data.frame(tapply(dat1$tabove30, list(dat1$admin_1, dat1$Epiyear), mean))
colnames(thot) <- c("th10", "th11", "th12", "th13", "tc14", "th15", "th16")
tcold <- as.data.frame(tapply(dat1$tbelow20, list(dat1$admin_1, dat1$Epiyear), mean))
colnames(tcold) <- c("tc10", "tc11", "tc12", "tc13", "tc14", "tc15", "tc16")



allStats <- as.data.frame(cbind(cpt, rainTot, tavg, thot, tcold))
#rownames(allStats) #notice that the row names are the province names
#colnames(allStats)


allStats <- allStats[-6,]

row.names(allStats) <- c("Cabo Delgado", "Gaza", "Inhambane", "Manica", "Maputo", "Nampula", "Nassa", "Sofala", "Tete", "Zambezia")

provpolydat <- SpatialPolygonsDataFrame(polyprov, allStats)

##### Now let's combine the `allStats` dataframe we created above with the `poly2` shapefile so that we can plot the statistics on the map.


```

```{r}
par(mfrow=c(2,1))
# MULTIPLE #
tempPal <- brewer.pal(n = 7, name = "YlOrRd")
rainPal <- brewer.pal(n = 7, name = "YlGnBu")
coldPal <- brewer.pal(n = 6, name = "Blues")
hotPal <- brewer.pal(n = 6, name = "Oranges")


trellis.par.set(sp.theme(regions=list(col = rainPal)))

p2 <- spplot(provpolydat, c("cpt11", "cpt12", "cpt13", "cpt14", "cpt15", "cpt16"), 
       names.attr = c("2011", "2012", "2013", "2014", "2015", "2016"), 
       colorkey=list(space="right"), scales = list(draw = FALSE), 
       main = "Total under 5 malaria cases by year", 
       as.table = TRUE, col.regions = tempPal, col='transparent', cuts=5)
#invisible(dev.off())
p2


p1 <- spplot(provpolydat, c("rain11", "rain12", "rain13", "rain14", "rain15", "rain16"),
       names.attr = c("2011", "2012", "2013", "2014", "2015", "2016"), 
       colorkey=list(space="right"), scales = list(draw = FALSE), 
       main = "Total rainfall by year", 
       as.table = TRUE, col.regions = rainPal, col='transparent',  cuts=5)
#invisible(dev.off())
p1

p3 <- spplot(provpolydat, c("tc10", "tc11", "tc12", "tc13", "tc14", "tc15", "tc16"),
       names.attr = c("2010", "2011", "2012", "2013", "2014", "2015", "2016"), 
       colorkey=list(space="right"), scales = list(draw = FALSE), 
       main = "Frequency of Colder Temperatures by year", 
       as.table = TRUE, col.regions = coldPal , col='transparent',  cuts=5)
 p3
 
 p4 <- spplot(provpolydat, c("th10", "th11", "th12", "th13", "tc14", "th15", "th16"),
       names.attr = c("2010", "2011", "2012", "2013", "2014", "2015", "2016"), 
       colorkey=list(space="right"), scales = list(draw = FALSE), 
       main = "Frequency of Hotter Temperatures by year", 
       as.table = TRUE, col.regions = tempPal, col='transparent',  cuts=5)
 p4
#invisible(dev.off())


```



BOXPLOTS

```{r}

mcpt <- ggplot(dat1, aes(Region, cpt, colour = as.factor(Epiyear))) + 
  geom_boxplot() +
  ggtitle(" Malaria Cases in under-5 Population")+
  scale_color_brewer(palette="Purples")+
  xlab("Region")+
  ylab("Cases per Thousand")+ 
  labs(colour="Year")+
  theme_dark()


raint <- ggplot(dat1, aes(Region, rainTot_log, colour = as.factor(Epiyear))) + 
  geom_boxplot() +
  ggtitle(" Average Log Weekly Rain by Region")+
  scale_color_brewer(palette="Purples")+
  xlab("Region")+
  ylab("Log(Weekly Rain Average)")+ 
  labs(colour="Year")+
  theme_dark()



tempt <- ggplot(dat1, aes(Region, tavg, colour = as.factor(Epiyear))) + 
  geom_boxplot() +
  ggtitle(" Average Weekly Temperature by Region")+
  scale_color_brewer(palette="Purples")+
  xlab("Region")+
  ylab("Weekly Average Temperature")+ 
  labs(colour="Year")+
  theme_dark()


rht <- ggplot(dat1, aes(Region, rh, colour = as.factor(Epiyear))) + 
  geom_boxplot() +
  ggtitle(" Average Humidity by Region")+
  scale_color_brewer(palette="Purples")+
  xlab("Region")+
  ylab("% Relative Humidity")+ 
  labs(colour="Year")+
  theme_dark()

raint
mcpt
tempt
rht

```

```{r, echo=FALSE}

dat1$Region_2 <- dat1$Region
revalue(dat1$Region_2, c("Center"="CE","Coastal"="CO","Northern"="NO", "Southern"="SO"))
dat1$Region_2 <- as.numeric(dat1$Region_2)
```


3D Plots
This plot is intended to three dimensionally show where cases are most elevated in relation to rain and temperature respectively lagged
```{r}


#scatterplot3js(x=dat1$rainTotlag8,y=dat1$tavglag8,z=dat1$cpt,size=0.4,color=as.factor(dat1$Region_2),bg="white")


plot3d(dat1$rainTotlag8,dat1$tavglag8,dat1$cpt, xlab="Rain Lag 8", ylab= "Temp Lag 8", zlab= "Cases per Thousand", col=as.factor(dat1$Region_2))


```


CUMULATIVE PLOTS

```{r}
dat9 <- subset(dat1, Epiyear >2010)
dat11 <- subset(dat9, Epiyear <= 2016)
### RAINFALL

ggplot(data = dat11)  +
  geom_smooth(mapping = aes(x = all.date, y = rainTot), color="black") +
  geom_smooth(mapping = aes(x = all.date, y = rainTotlag8), color="red") +
  geom_smooth(mapping = aes(x = all.date, y = rainTotlag6), color="pink") +
  geom_smooth(mapping = aes(x = all.date, y = rainTotlag4), color="orange")+
  labs(x = "Epidemiology week", y = "Rainfall")

### Temperature
ggplot(data = dat11) + 
  geom_smooth(mapping = aes(x = all.date, y = tavg), color="black") +
  geom_smooth(mapping = aes(x = all.date, y = tavglag8), color="red") +
  geom_smooth(mapping = aes(x = all.date, y = tavglag6), color="pink") +
  geom_smooth(mapping = aes(x = all.date, y = tavglag4), color="orange")+
  labs(x = "Epidemiology week", y = "Temperature")



```
```{r}

dat10 <- subset(dat1, Epiyear > 2010)
### Temperature
ggplot(data = dat10) + 
  geom_smooth(mapping = aes(x = all.date, y = cpt), color="black") +
  geom_smooth(mapping = aes(x = all.date, y = tavglag8), color="orange") +
  geom_smooth(mapping = aes(x = all.date, y = rainTotlag8), color="turquoise")+
  scale_y_continuous(sec.axis = sec_axis(~.*2, name = "Weather Patterns")) +
  labs(x = "Time", y = "Cases per 1,000")




ggplot(data = dat10) + 
  geom_smooth(mapping = aes(x = all.date, y = cpt), color="black") +
  geom_smooth(mapping = aes(x = all.date, y = rainTotlag8), color="turquoise")+
  labs(x = "Time", y = "Cases per 1,000")
```
par (mfrow=c(2,2))

plot(dat1$rainTot, dat1$cpt)
plot(dat1$rainTotlag2, dat1$cpt)
plot(dat1$rainTotlag4, dat1$cpt)
plot(dat1$rainTotlag8, dat1$cpt)
```