---
title: "RPROJECT"
author: "Melissa Lowe"
date: "10/22/2018"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(plyr)
library(dplyr)
library(ggplot2)
library(lattice)
library(tidyverse)
library(modelr)
library(RColorBrewer)
library(sp)
library(maptools) 
library(lattice)
library(latticeExtra) # For layer()
library(rgdal)
library(classInt)
library(gridExtra)
library(grid)
library(knitr)
library(tmap)
library(tmaptools)
library(plot3D)
library(rgl)
library(threejs)

```
Shape Files:

```{r, include=FALSE}
# read in the Moz shape file for districts
polydist <- readShapePoly('~/Desktop/MS YEAR 2/S1/R Class/Moz_admin2.shp', IDvar="DISTCODE")
#this is the polygon file. it's been developed with district codes. RStudio helps recall how to get things out of lists. They're shape files as R stores them
#plot(polydist)

polyprov <- readShapePoly('~/Desktop/MS YEAR 2/S1/R Class/mozambique_admin1.shp', IDvar="admin_1")
#this is the polygon file. it's been developed with province codes. RStudio helps recall how to get things out of lists. They're shape files as R stores them

#plot(polyprov)
```


Data Steps:

```{r, include=FALSE}
dat10 <- read.csv("~/Desktop/MS YEAR 2/S1/R Class/MozSyntheticMalaria.csv")

dat10$rainTotlag2 <- lag(dat10$rainTot, 2)

dat10$rainTotlag4 <- lag(dat10$rainTot, 4)

dat10$rainTotlag6 <- lag(dat10$rainTot, 6)

dat10$rainTotlag8<- lag(dat10$rainTot, 8)


dat10$tavglag2 <- lag(dat10$tavg, 2)

dat10$tavglag4 <- lag(dat10$tavg, 4)

dat10$tavglag6 <- lag(dat10$tavg, 6)

dat10$tavglag8<- lag(dat10$tavg, 8)

dat10$rhlag2 <- lag(dat10$rh, 2)

dat10$rhlag4 <- lag(dat10$rh, 4)

dat10$rhlag6 <- lag(dat10$rh, 6)

dat10$rhlag8 <- lag(dat10$rh, 8)

dat10$cpt <- (dat10$malaria/(dat10$Population_UN*dat10$u5weight))*1000

#proportion of under-5 population that had malaria each week

dat10 <- mutate(dat10, all.date = as.Date(paste(dat10$Epiyear, dat10$Epiweek, "1", sep="-"), format="%Y-%U-%u")) #create a longterm date 

dat1 <- subset(dat10, Epiyear <2017)

#made low values of rainTot 1 so that a logged value of it is 0 on the low end to show range of total rain
dat1$rainTot_log <- ifelse(dat1$rainTot < 1, 1, log(dat1$rainTot))

#What about the density of malaria cases in each district? Ie. Malaria cases per 1000 square km?

#malaria is the number of cases under 5 reported that week (this is simulated)
dat1$malsqkm <- (dat1$malaria/dat1$SQKM)*1000

range(dat1$SQKM)

range(dat1$malsqkm)

range(dat1$rainTot)

range(dat1$rainTotlag8)
```

```{r, include=FALSE}
par(mfrow=c(1,2))
hist(dat1$rainTot)
hist(dat1$rainTot_log)
range(dat1$rainTot_log)
```

CDC. “PMI, Mozambique Profile.” President’s Malaria Initiative, USAID, 2018, www.pmi.gov/where-we-work/mozambique.
President’s Malaria Initiative. “Mozambique, Malaria Operational Plan FY 2018.” President’s Malaria Initiative, USAID, 2018, www.pmi.gov/docs/default-source/default-document-library/malaria-operational-plans/fy-2018/fy-2018-mozambique-malaria-operational-plan.pdf?sfvrsn=5.


```{r, echo=FALSE}

tab1 <- dat1 %>% 
  group_by(Epiyear) %>% 
  summarise(
  malaria_m = round(mean(malaria, na.rm = TRUE),2),
  rainTot_med = round(median(rainTot, na.rm = TRUE),2),
  tavg_m = round(mean(tavg, na.rm = TRUE),2),
  rh = round(mean(rh, na.rm = TRUE),2)
) 
#kable(tab1) #this is now a data frame. we can edit it to make it prettier


# Let's make it more attractive
colnames(tab1) <- c("Year", "Malaria (cases)", "Median Rainfall (mm)", "Average Temp(C)", "Relative Humidity(%)")
kable(tab1)


```


DISTRICT BASED SPATIAL MAPPING


```{r, include=FALSE}
#as.factor(dat1$SQKM) # there are 142 levels in square km which means they are by district
#dat1 <- subset(dat1, dat1$Province != "MAPUTO CIDADE")

dat1$admin_2 <- dat1$DISTCODE

#want to know statistics for district code and year, rows are distcode and columns are years
cpt <- as.data.frame(tapply(dat1$cpt,list(dat1$admin_2, dat1$Epiyear), max))
colnames(cpt) <- c("cpt10", "cpt11", "cpt12", "cpt13", "cpt14", "cpt15", "cpt16")
rainTot <- as.data.frame(tapply(dat1$rainTot, list(dat1$admin_2, dat1$Epiyear), sum))
colnames(rainTot) <- c("rain10", "rain11", "rain12", "rain13", "rain14", "rain15", "rain16")
tavg1 <- as.data.frame(tapply(dat1$tavg, list(dat1$admin_2, dat1$Epiyear), mean))
colnames(tavg1) <- c("t10", "t11", "t12", "t13", "t14", "t15", "t16")
rhavg <- as.data.frame(tapply(dat1$rh, list(dat1$admin_2, dat1$Epiyear), mean))
colnames(rhavg) <- c("rh10", "rh11", "rh12", "rh13", "rh14", "rh15", "rh16")



allStats2 <- as.data.frame(cbind(cpt, rainTot, tavg1, rhavg))
#rownames(allStats) #notice that the row names are the province names
colnames(allStats2)

#allStats2$mean_cpt <- allStats2 %>% mutate_at(vars(matches("cpt")), mean)

#rownames(allStats2)

distpolydat <- SpatialPolygonsDataFrame(polydist, allStats2)
```

I am going to focus on Epiyear 2016 because that's where we appear to have the highest malaria, rainfall, and temperatures - this could also be because of more data within that year

The first epi week of the year ends, by definition, on the first Saturday of January, as long as it falls at least four days into the month. Each epi week begins on a Sunday and ends on a Saturday.

Palettes
```{r, include=FALSE}
tempPal <- brewer.pal(n = 9, name = "YlOrRd")
rainPal <- brewer.pal(n = 9, name = "YlGnBu")
coldPal <- brewer.pal(n = 9, name = "Blues")
hotPal <- brewer.pal(n = 9, name = "Oranges")
cptPal <-brewer.pal(n = 9, name = "PuRd") 
```

```{r, echo=FALSE}

rain6 <- spplot(distpolydat, "rain16", col.regions = rainPal, cuts = 5, col = "transparent", 
       main = "Tot.Rainfall (mm)", sub = "By District")
cpt6 <- spplot(distpolydat, "cpt16", col.regions = cptPal, cuts = 5, col = "transparent", 
       main = "Max Cases/1000 \n Under-5", sub = "By District")
temp6 <- spplot(distpolydat, "t16", col.regions = tempPal, cuts = 5, col = "transparent", 
       main = "Ave.Temperature\n (Celsius)", sub = "By District")

grid.arrange(rain6, cpt6,temp6, ncol = 3, top= "2016 Mozambique Malaria Data")
```
2016
```{r, echo=FALSE}
tm_shape(distpolydat) +
tm_polygons("rain16", palette = rainPal, title = "2016 Total Rainfall", contrast = 0.7, border.col = "transparent")+
tm_squares("cpt16", col = "t16", palette = "-RdYlGn", breaks = c(15, 17, 19, 21, 23, 25, 27), border.col="transparent",
title.size = "Max Cases per Thousand (2016)", title.col = "2016 Average Temperature") +
tm_legend(outside = TRUE, bg.color = "mistyrose2", frame = TRUE)+
tm_layout(title = "Mozambique Malaria Summary in 2016",
title.position = c("center", "top"), 
legend.position = c("right", "bottom"), frame = FALSE,
inner.margins = c(0.1, 0.1, 0.05, 0.05), bg.color="mistyrose4")
```

2015
```{r, echo=FALSE}
tm_shape(distpolydat) +
tm_polygons("rain15", palette = rainPal, title = "2015 Total Rainfall", contrast = 0.7, border.col = "transparent")+
tm_squares("cpt16", col = "t15", palette = "-RdYlGn", breaks = c(15, 17, 19, 21, 23, 25, 27), border.col="transparent",
title.size = "Max Cases per Thousand (2015)", title.col = "2015 Average Temperature") +
tm_legend(outside = TRUE, bg.color = "mistyrose2", frame = TRUE)+
tm_layout(title = "Mozambique Malaria Summary in 2015",
  title.position = c("center", "top"), 
  legend.position = c("right", "bottom"), frame = FALSE,
  inner.margins = c(0.1, 0.1, 0.05, 0.05), bg.color="mistyrose4")
```
2014
```{r, echo=FALSE}
tm_shape(distpolydat) +
tm_polygons("rain14", palette = rainPal, title = "Total 2014 Rainfall", contrast = 0.7, border.col = "transparent")+
tm_squares("cpt14", col = "t14", palette = "-RdYlGn", breaks = c(15, 17, 19, 21, 23, 25, 27), border.col="transparent",
title.size = "Max Cases per Thousand (2014)", title.col = "2014 Average Temperature") +
tm_legend(outside = TRUE, bg.color = "mistyrose2", frame = TRUE)+
tm_layout(title = "Mozambique Malaria Summary in 2014",
  title.position = c("center", "top"), 
  legend.position = c("right", "bottom"), frame = FALSE,
  inner.margins = c(0.1, 0.1, 0.05, 0.05), bg.color="mistyrose4")
```
2013
```{r, echo=FALSE}
tm_shape(distpolydat) +
  tm_polygons("rain13", palette = rainPal, title = "2013 Total Rainfall", contrast = 0.7, border.col = "transparent")+
tm_squares("cpt13", col = "t13", palette = "-RdYlGn", breaks = c(15, 17, 19, 21, 23, 25, 27), border.col="transparent",
title.size = "Max Cases per Thousand (2013)", title.col = "2013 Average Temperature") +
tm_legend(outside = TRUE, bg.color = "mistyrose2", frame = TRUE)+
tm_layout(title = "Mozambique Malaria Summary in 2013",
  title.position = c("center", "top"), 
  legend.position = c("right", "bottom"), frame = FALSE,
  inner.margins = c(0.1, 0.1, 0.05, 0.05), bg.color="mistyrose4")
```
2012
```{r, echo=FALSE}
tm_shape(distpolydat) +
tm_polygons("rain12", palette = rainPal, title = "2012 Total Rainfall", contrast = 0.7, border.col = "transparent")+
tm_squares("cpt12", col = "t12", palette = "-RdYlGn", breaks = c(15, 17, 19, 21, 23, 25, 27), border.col="transparent",
title.size = "Max Cases per Thousand (2012)", title.col = "2012 Average Temperature") +
tm_legend(outside = TRUE, bg.color = "mistyrose2", frame = TRUE)+
tm_layout(title = "Mozambique Malaria Summary in 2012",
  title.position = c("center", "top"), 
  legend.position = c("right", "bottom"), frame = FALSE,
  inner.margins = c(0.1, 0.1, 0.05, 0.05), bg.color="mistyrose4")

```
2011
```{r, echo=FALSE}
tm_shape(distpolydat) +
tm_polygons("rain11", palette = rainPal, title = "2011 Total Rainfall", contrast = 0.7, border.col = "transparent")+
tm_squares("cpt11", col = "t11", palette = "-RdYlGn", breaks = c(15, 17, 19, 21, 23, 25, 27), border.col="transparent",
title.size = "Max Cases per Thousand (2011)", title.col = "2011 Average Temperature") +
tm_legend(outside = TRUE, bg.color = "mistyrose2", frame = TRUE)+
tm_layout(title = "Mozambique Malaria Summary in 2011",
  title.position = c("center", "top"), 
  legend.position = c("right", "bottom"), frame = FALSE,
  inner.margins = c(0.1, 0.1, 0.05, 0.05), bg.color="mistyrose4")

```
2010
```{r, echo=FALSE}
tm_shape(distpolydat) +
tm_polygons("rain10", palette = rainPal, title = "2010 Total Rainfall", contrast = 0.7, border.col = "transparent")+
tm_squares("cpt11", col = "t10", palette = "-RdYlGn", breaks = c(15, 17, 19, 21, 23, 25, 27), border.col="transparent",
title.size = "Max Cases per Thousand (2010)", title.col = "2010 Average Temperature") +
tm_legend(outside = TRUE, bg.color = "mistyrose2", frame = TRUE)+
tm_layout(title = "Mozambique Malaria Summary in 2010",
  title.position = c("center", "top"), 
  legend.position = c("right", "bottom"), frame = FALSE,
  inner.margins = c(0.1, 0.1, 0.05, 0.05), bg.color="mistyrose4")


```


BOXPLOTS

```{r, echo=FALSE}

mcpt <- ggplot(dat1, aes(Region, cpt, colour = as.factor(Epiyear))) + 
  geom_boxplot() +
  ggtitle(" Malaria Cases in under-5 Population")+
  scale_color_brewer(palette="Purples")+
  xlab("Region")+
  ylab("Cases per Thousand")+ 
  labs(colour="Year")+
  theme_dark()


raint <- ggplot(dat1, aes(Region, rainTot_log, colour = as.factor(Epiyear))) + 
  geom_boxplot() +
  ggtitle(" Average Log Weekly Rain by Region")+
  scale_color_brewer(palette="Purples")+
  xlab("Region")+
  ylab("Log(Weekly Rain Average)")+ 
  labs(colour="Year")+
  theme_dark()



tempt <- ggplot(dat1, aes(Region, tavg, colour = as.factor(Epiyear))) + 
  geom_boxplot() +
  ggtitle(" Average Weekly Temperature by Region")+
  scale_color_brewer(palette="Purples")+
  xlab("Region")+
  ylab("Weekly Average Temperature")+ 
  labs(colour="Year")+
  theme_dark()


rht <- ggplot(dat1, aes(Region, rh, colour = as.factor(Epiyear))) + 
  geom_boxplot() +
  ggtitle(" Average Humidity by Region")+
  scale_color_brewer(palette="Purples")+
  xlab("Region")+
  ylab("% Relative Humidity")+ 
  labs(colour="Year")+
  theme_dark()

raint
mcpt
tempt
rht

```



CUMULATIVE PLOTS

```{r, echo=FALSE}
dat9 <- subset(dat1, Epiyear >2010)
dat11 <- subset(dat9, Epiyear <= 2016)
### RAINFALL

ggplot(data = dat11)  +
  geom_smooth(mapping = aes(x = all.date, y = rainTot), color="black") +
  geom_smooth(mapping = aes(x = all.date, y = rainTotlag8), color="red") +
  geom_smooth(mapping = aes(x = all.date, y = rainTotlag6), color="pink") +
  geom_smooth(mapping = aes(x = all.date, y = rainTotlag4), color="orange")+ 
  geom_smooth(mapping = aes(x = all.date, y = cpt), color="turquoise")+
  labs(x = "Epidemiology week", y = "Rainfall")

### Temperature
ggplot(data = dat11) + 
  geom_smooth(mapping = aes(x = all.date, y = tavg), color="black") +
  geom_smooth(mapping = aes(x = all.date, y = tavglag8), color="red") +
  geom_smooth(mapping = aes(x = all.date, y = tavglag6), color="pink") +
  geom_smooth(mapping = aes(x = all.date, y = tavglag4), color="orange")+
  geom_smooth(mapping = aes(x = all.date, y = cpt), color="turquoise")+
  labs(x = "Epidemiology week", y = "Temperature")

#Humidity

ggplot(data = dat11) + 
  geom_smooth(mapping = aes(x = all.date, y = rh), color="black") +
  geom_smooth(mapping = aes(x = all.date, y = rhlag8), color="red") +
  geom_smooth(mapping = aes(x = all.date, y = rhlag6), color="pink") +
  geom_smooth(mapping = aes(x = all.date, y = rhlag4), color="orange")+
  geom_smooth(mapping = aes(x = all.date, y = cpt), color="turquoise")+
  labs(x = "Epidemiology week", y = "Relative Humidity")

```



