---
title: "map hw"
author: "Melissa Lowe"
date: "10/25/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

 #### **Maps**

```{r, echo=F, message=F, warning=F}

op <- par()
options(width=80)
emptyenv()
rm(list=ls())

### MAPPING PACKAGES ###
#install.packages(c("RColorBrewer", "sp", "maptools", "lattice", "latticeExtra", "rgdal"))
library(RColorBrewer)
library(sp)
library(maptools) 
library(lattice)
library(latticeExtra) # For layer()
library(rgdal)
library(RColorBrewer)
library(classInt)
library(gridExtra)
library(grid)
```

##### Now we need to aggregate the malaria and climatic data over the districts to create a set of statistics for each district. Conveniently, this aggregation I perform results in the rownames of the dataframe being the district codes, which is required for creating what is called a spatial polygons dataframe (we'll create this below).
```{r, echo=FALSE, message=T, warning=T}
# Note: because I started an Rproject, I can use relative files paths
# I just need to use './...' to access other folders within this project now
all2 <- read.csv("~/Desktop/MS YEAR 2/S1/R Class/R Class Notes/R-Class/MozSyntheticMalaria.csv")
all2$cpt <- (all2$malaria/(all2$Population_UN*all2$u5weight))*1000
all2.2 <- subset(all2, Epiyear < 2017)

```

```{r, echo=FALSE}
table(all2.2$Province)

all2.2 <- subset(all2.2, all2.2$Province != "MAPUTO CIDADE")

all2.2$admin_1 <- all2.2$Province

# Create means by province and epiyear
malProv <- tapply(all2.2$malaria, list(all2.2$Province, all2.2$Epiyear), mean)
rainProv <- tapply(all2.2$rainTot, list(all2.2$Province, all2.2$Epiyear), mean)

```

```{r, echo=FALSE}

#want to know statistics for district code and year, rows are distcode and columns are years
cpt <- as.data.frame(tapply(all2.2$malaria, list(all2.2$admin_1, all2.2$Epiyear), sum))
colnames(cpt) <- c("cpt10", "cpt11", "cpt12", "cpt13", "cpt14", "cpt15", "cpt16")
rainTot <- as.data.frame(tapply(all2.2$rainTot, list(all2.2$admin_1, all2.2$Epiyear), sum))
colnames(rainTot) <- c("rain10", "rain11", "rain12", "rain13", "rain14", "rain15", "rain16")
tavg <- as.data.frame(tapply(all2.2$tavg, list(all2.2$admin_1, all2.2$Epiyear), mean))
colnames(tavg) <- c("t10", "t11", "t12", "t13", "t14", "t15", "t16")
  
allStats <- as.data.frame(cbind(cpt, rainTot, tavg))
#rownames(allStats) #notice that the row names are the province names
#colnames(allStats)

allStats <- allStats[-6,]
```

```{r, echo=FALSE}
##### Let's load the district-level administrative shape file for Mozambique.

# read in the Moz shape file for districts
poly1 <- readShapePoly('~/Desktop/MS YEAR 2/S1/R Class/R Class Notes/R-Class/Moz_admin2.shp', IDvar="DISTCODE")
#this is the polygon file. it's been developed with district codes. RStudio helps recall how to get things out of lists. They're shape files as R stores them


poly2 <- readShapePoly('~/Desktop/MS YEAR 2/S1/R Class/R Class Notes/R-Class/mozambique_admin1.shp', IDvar="admin_1")
#this is the polygon file. it's been developed with province codes. RStudio helps recall how to get things out of lists. They're shape files as R stores them



row.names(allStats) <- c("Cabo Delgado", "Gaza", "Inhambane", "Manica", "Maputo", "Nampula", "Nassa", "Sofala", "Tete", "Zambezia")


##### Now let's combine the `allStats` dataframe we created above with the `poly2` shapefile so that we can plot the statistics on the map.

polydat <- SpatialPolygonsDataFrame(poly2, allStats)
```


this code is paritally from class, partially from Jimmy and partially from Emma: (Thanks peeps)
```{r, echo=F, message=T, warning=T, fig.height=8, fig.width=8}
par(mfrow=c(2,1))
# MULTIPLE #
tempPal <- brewer.pal(n = 7, name = "YlOrRd")
rainPal <- brewer.pal(n = 7, name = "YlGnBu")
my.palette <- brewer.pal(n = 7, name = "OrRd")


trellis.par.set(sp.theme(regions=list(col = rainPal)))

p2 <- spplot(polydat, c("cpt11", "cpt12", "cpt13", "cpt14", "cpt15", "cpt16"), 
       names.attr = c("2011", "2012", "2013", "2014", "2015", "2016"), 
       colorkey=list(space="right"), scales = list(draw = TRUE), 
       main = "Total under 5 malaria cases by year", 
       as.table = TRUE, col.regions = tempPal, col='transparent', cuts=5)
#invisible(dev.off())



p1 <- spplot(polydat, c("rain11", "rain12", "rain13", "rain14", "rain15", "rain16"),
       names.attr = c("2011", "2012", "2013", "2014", "2015", "2016"), 
       colorkey=list(space="right"), scales = list(draw = TRUE), 
       main = "Total rainfall by year", 
       as.table = TRUE, col.regions = rainPal, col='transparent',  cuts=5)
#invisible(dev.off())
trellis.par.set(sp.theme(regions=list(col = rainPal)))
grid.arrange(p1, p2, ncol = 2)
```


